Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-03-16 Grassmannian_cells.tex
small changes

If each component $C_i$ admits a decomposition into affine spaces, Theorem~\ref{thm:bb} implies the same is true of $X$.
Indeed, we can trivialize each affine bundle $\Att(C_i)\to C_i$ over each affine piece of $C_i$ to get a decomposition of $X$ into affine spaces.

@@

In the following, we say that $d:\supp(\hat X)\to\ZZ$ satisfies the \emph{degree condition} for $\hat X$ if it has the properties of Lemma~\ref{le:degree condition}.
\begin{theorem}\label{thm:torusfixedpoints}
  Let $X$ be a representation of $Q$ which can be lifted to a representation $\hat X$ of $\hat Q$ and choose $d:\supp(\hat X)\to\ZZ$ such that it satisfies the degree condition for $\hat X$.
  Then the $\CC^*$-action on $\bigoplus_{i\in Q_0} X_i$ determined by $t.x_{(i,\chi)}=t^{d(i,\chi)}x_{(i,\chi)}$ for $x_{(i,\chi)}\in X_{(i,\chi)}$ induces a $\CC^*$-action on $\Gr_\bfe^Q(X)$ such that
  \[\Gr^Q_\bfe(X)^{\CC^*}\cong\bigsqcup_{\hat\bfe} \Gr^{\hat Q}_{\hat\bfe}(\hat X),\]
  where $\hat\bfe$ runs through all dimension vectors compatible with $\bfe$.
\end{theorem}
\begin{proof}
  A representation $U\in\Gr_{\bfe}(X)$ is a $\CC^*$ fixed point if and only if $t.U=U$ for all $t\in\C^\ast$, i.e.\ $t.U_i=U_i$ for all $i\in Q_0$ and all $t\in\C^\ast$.
  Thus, apart from being a subrepresentation of $X$, each component $U_i$ is a fixed point of the induced $\CC^*$-actions on the usual Grassmannians of vector subspaces $\Gr_{\bfe_i}(X_i)$.
  By Lemma~\ref{le:usualGrass}, this holds precisely when we have a decomposition 
  \[U_i=\bigoplus_{\chi\in A_Q} U_i\cap X_{(i,\chi)}\]
  which is equivalent to $U$ being liftable to the universal abelian covering $\hat Q$.
\end{proof}
The next step is to iterate the $\CC^*$-actions, keeping in mind the following idea: every representation $X$ which lifts to the universal covering quiver also lifts to the universal abelian covering quiver and to the iterated universal abelian covering quivers, i.e.\ to each $\hat Q^{(k)}:=\widehat{\hat Q^{(k-1)}}$ with $\hat Q^{(1)}:=\hat Q$.
Now it is straightforward to check that there exist natural surjective morphisms $f_k:\widetilde Q\to \hat Q^{(k)}$ which become injective on finite subquivers if $k\gg 0$, see also \cite[Section 3.4]{wei}.
Since the support of $X$ is finite as a representation of $\tilde Q$, we can find $k\geq 0$ such that the full subquiver with vertices $\supp(X)\subseteq \hat Q^{(k+1)}_0$ is a tree.
Thus, writing $\hat X^{(\ell)}$ for the lift of $X$ to $\hat Q^{(\ell)}$, there exists a $\CC^*$-action on the vector spaces $\hat X^{(k)}_\beta$ for $\beta\in \hat Q^{(k-1)}_0\times A_{\hat Q^{(k-1)}_1}$ which induces $\CC^*$-actions on the quiver Grassmannians $\Gr_{\hat\bfe^{(k)}}^{\hat Q^{(k)}}\big(\hat X^{(k)}\big)$ such that the fixed point sets are precisely $\Gr_{\hat\bfe^{(k+1)}}^{\hat Q^{(k+1)}}\big(\hat X^{(k+1)}\big)$.
If we denote these iterated torus fixed points by $\Gr^Q_\bfe(X)^{(k+1)}$, we obtain the following result.
\begin{corollary}
  Let $X$ be a representation which can be lifted to $\tilde Q$.
  Then there exists an iterated torus action such that
  \[\Gr^Q_\bfe(X)^{(k+1)}\cong \bigsqcup_{\hat\bfe^{(k)}} \Gr_{\hat\bfe^{(k)}}^{\hat Q^{(k)}}\big(\hat X^{(k)}\big)^{\CC^*}\cong \bigsqcup_{\hat\bfe^{(k+1)}} \Gr^{\hat Q^{(k+1)}}_{\hat\bfe^{(k+1)}}\big(\hat X^{(k+1)}\big)\cong \bigsqcup_{\tilde\bfe} \Gr^{\tilde Q}_{\tilde\bfe}(\tilde X),\]
  where $\hat\bfe^{(k)},\,\hat\bfe^{(k+1)},\,\tilde\bfe$ run through all dimension vectors compatible with $\bfe$.
\end{corollary}
%\begin{corollary}
%  affine bundles over $\Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)$ if $\Gr^Q_\bfe(X)$ is smooth
%  \[\{U\in \Gr^Q_\bfe(X):\lim_{t\to0} t\cdot U\in \Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)\}\]
%\end{corollary}
Define the $F$-polynomial of a representation $X$ by 
$$F_X=\sum_{\bfe\in\NN^{Q_0}}\chi(\Gr_\bfe(X))y^\bfe\in\ZZ[y_i\mid i\in Q_0].$$

@@

Let $\cG_{\tbff,\tbfg}:=\Psi^{-1}_\tbfe\big(\Gr_\tbff(M)\times \Gr_\tbfg(N)\big)$. Then we have
$$\Gr_\tbfe(B)=\bigsqcup_{\tbff+\tbfg=\tbfe} \cG_{\tbff,\tbfg}.$$
The map $\Psi_\tbfe$ restricts to morphisms of projective varieties
\[\Psi_\tbfe^{\tbff,\tbfg}:\cG_{\tbff,\tbfg}\to \Gr_\tbff(M)\times \Gr_\tbfg(N).\]
The following is proved in \cite[Section 3]{cefr}.
\begin{theorem}
  \label{vb}
  If the fiber dimension of $\Psi_\tbfe^{\tbff,\tbfg}$ is constant over $\mathrm{Im}\big(\Psi_\tbfe^{\tbff,\tbfg}\big)$, then it is an affine bundle over the image.
  In particular, the existence of a cell decomposition of $\mathrm{Im}\big(\Psi_\tbfe^{\tbff,\tbfg}\big)$ implies a cell decomposition of $\cG_{\tbff,\tbfg}$ in this case.
\end{theorem}
%\begin{proof}\footnote{sketch, double-check}
%As $X$ and $Y$ are projective, they are compact. As the fibers are affine spaces of constant dimension and as $X$ and $Y$ are smooth, the induced morphism on the tangent spaces is surjective for all points $x\in X$. This shows that $f$ is smooth of relative dimension $n$, see Hartshorne and thus a submersion?!
%\end{proof}

>>>>

If each component $C_i$ admits a cell decomposition, Theorem~\ref{thm:bb} implies the same is true of $X$. Indeed, we can trivialize each affine bundle $\Att(C_i)\to C_i$ over each affine piece of $C_i$ and then taking the natural ordering of the resulting affine spaces gives a cell decomposition of $X$.

@@

In the following, we say that $d:\supp(\hat X)\to\ZZ$ satisfies the \emph{degree condition} for $\hat X$ if it has the properties of Lemma~\ref{le:degree condition}.
\begin{theorem}\label{thm:torusfixedpoints}
  Let $X$ be a representation of $Q$ which can be lifted to a representation $\hat X$ of $\hat Q$ and choose $d:\supp(\hat X)\to\ZZ$ such that it satisfies the degree condition for $\hat X$.
  Then the $\CC^*$-action on $\bigoplus_{i\in Q_0} X_i$ determined by $t.x_{(i,\chi)}=t^{d(i,\chi)}x_{(i,\chi)}$ for $x_{(i,\chi)}\in X_{(i,\chi)}$ induces a $\CC^*$-action on $\Gr_\bfe^Q(X)$ such that
  \[\Gr^Q_\bfe(X)^{\CC^*}\cong\bigsqcup_{\hat\bfe} \Gr^{\hat Q}_{\hat\bfe}(\hat X),\]
  where $\hat\bfe$ runs through all dimension vectors compatible with $\bfe$.
\end{theorem}
\begin{proof}
  A representation $U\in\Gr_{\bfe}(X)$ is a $\CC^*$ fixed point if and only if $t.U=U$ for all $t\in\C^\ast$, i.e.\ $t.U_i=U_i$ for all $i\in Q_0$ and all $t\in\C^\ast$.
  Thus, apart from being a subrepresentation of $X$, each component $U_i$ is a fixed point of the induced $\CC^*$-actions on the usual Grassmannians of vector subspaces $\Gr_{\bfe_i}(X_i)$.
  By Lemma~\ref{le:usualGrass}, this holds precisely when we have a decomposition 
  \[U_i=\bigoplus_{\chi\in A_Q} U_i\cap X_{(i,\chi)}\]
  which is equivalent to $U$ being liftable to the universal abelian covering $\hat Q$.
\end{proof}
The next step is to iterate the $\CC^*$-actions, keeping in mind the following idea: every representation $X$ which lifts to the universal covering quiver also lifts to the universal abelian covering quiver and to the iterated universal abelian covering quivers, i.e.\ to each $\hat Q^{(k)}:=\widehat{\hat Q^{(k-1)}}$ with $\hat Q^{(1)}:=\hat Q$.
Now it is straightforward to check that there exist natural surjective morphisms $f_k:\widetilde Q\to \hat Q^{(k)}$ which become injective on finite subquivers if $k\gg 0$, see also \cite[Section 3.4]{wei}.
Since the support of $X$ is finite as a representation of $\tilde Q$, we can find $k\geq 0$ such that the full subquiver with vertices $\supp(X)\subseteq \hat Q^{(k+1)}_0$ is a tree.
Thus, writing $\hat X^{(\ell)}$ for the lift of $X$ to $\hat Q^{(\ell)}$, there exists a $\CC^*$-action on the vector spaces $\hat X^{(k)}_\beta$ for $\beta\in \hat Q^{(k-1)}_0\times A_{\hat Q^{(k-1)}}$ which induces $\CC^*$-actions on the quiver Grassmannians $\Gr_{\hat\bfe^{(k)}}^{\hat Q^{(k)}}\big(\hat X^{(k)}\big)$ such that the fixed point sets are precisely $\Gr_{\hat\bfe^{(k+1)}}^{\hat Q^{(k+1)}}\big(\hat X^{(k+1)}\big)$.
If we denote these iterated torus fixed points by $\Gr^Q_\bfe(X)^{(k+1)}$, we obtain the following result.
\begin{corollary}
  Let $X$ be a representation which can be lifted to $\tilde Q$.
  Then there exists an iterated torus action such that
  \[\Gr^Q_\bfe(X)^{(k+1)}\cong \bigsqcup_{\hat\bfe^{(k)}} \Gr_{\hat\bfe^{(k)}}^{\hat Q^{(k)}}\big(\hat X^{(k)}\big)^{\CC^*}\cong \bigsqcup_{\hat\bfe^{(k+1)}} \Gr^{\hat Q^{(k+1)}}_{\hat\bfe^{(k+1)}}\big(\hat X^{(k+1)}\big)\cong \bigsqcup_{\tilde\bfe} \Gr^{\tilde Q}_{\tilde\bfe}(\tilde X),\]
  where $\hat\bfe^{(k)},\,\hat\bfe^{(k+1)},\,\tilde\bfe$ run through all dimension vectors compatible with $\bfe$.
\end{corollary}
%\begin{corollary}
%  affine bundles over $\Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)$ if $\Gr^Q_\bfe(X)$ is smooth
%  \[\{U\in \Gr^Q_\bfe(X):\lim_{t\to0} t\cdot U\in \Gr^{\tilde Q}_{\tilde\bfe}(\tilde X)\}\]
%\end{corollary}
Define the $F$-polynomial of a representation $X$ by 
$$F_X=\sum_{\bfe\in\NN^{Q_0}}\chi(\Gr_\bfe(X))y^\bfe\in\ZZ[y_i\mid i\in Q_0].$$

@@

For $\cG_{\tbff,\tbfg}:=\Psi^{-1}_\tbfe\big(\Gr_\tbff(M)\times \Gr_\tbfg(N)\big)$, we have $$\Gr_\tbfe(B)=\bigsqcup_{\tbff+\tbfg=\tbfe} \cG_{\tbff,\tbfg}.$$
Then $\Psi_\tbfe$ restricts to a map
\[\Psi_\tbfe^{\tbff,\tbfg}:\cG_{\tbff,\tbfg}\to \Gr_\tbff(M)\times \Gr_\tbfg(N).\]
The following is proved in \cite[Section 3]{cefr}.
\begin{theorem}
  \label{vb}
  If $\mathrm{Im}\big(\Psi_\tbfe^{\tbff,\tbfg}\big)$ is locally closed and the fiber dimension of $\Psi_\tbfe^{\tbff,\tbfg}$ is constant over $\mathrm{Im}\big(\Psi_\tbfe^{\tbff,\tbfg}\big)$, then $\Psi_\tbfe^{\tbff,\tbfg}:\cG_{\tbff,\tbfg}\to \mathrm{Im}\big(\Psi_\tbfe^{\tbff,\tbfg}\big)$ is an affine bundle.
  In particular, the existence of a cell decomposition of $\mathrm{Im}\big(\Psi_\tbfe^{\tbff,\tbfg}\big)$ implies a cell decomposition of $\cG_{\tbff,\tbfg}$ in this case.
\end{theorem}
%\begin{proof}\footnote{sketch, double-check}
%As $X$ and $Y$ are projective, they are compact. As the fibers are affine spaces of constant dimension and as $X$ and $Y$ are smooth, the induced morphism on the tangent spaces is surjective for all points $x\in X$. This shows that $f$ is smooth of relative dimension $n$, see Hartshorne and thus a submersion?!
%\end{proof}
